---
title: "p8105_hw2_az2587"
author: "Anyu Zhu"
date: "9/30/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
```

## Problem 1

Read and clean data from Mr. Trash Wheel dataset:
```{r}
trash_wheel = read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                         sheet = "Mr. Trash Wheel",
                         range =  cell_cols("A:N"), col_names = TRUE, 
                         trim_ws = TRUE, skip = 1) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = round(sports_balls, 0))

trash_wheel
```

Read and clean precipitation data for 2018 and 2019:
```{r}
precipitation_2018 = read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                                sheet = "2018 Precipitation", range = "A2:B14", 
                                col_names = TRUE, trim_ws = TRUE) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2018) %>% 
  select(year, month, total)

precipitation_2019 = read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                                sheet = "2019 Precipitation", range = "A2:B14", 
                                col_names = TRUE, trim_ws = TRUE) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2019) %>% 
  select(year, month, total)

precipitation = 
  bind_rows(precipitation_2018, precipitation_2019) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month])

precipitation
```
**Conclusion**
```{r}
analysis_trash_wheel = 
  tibble(
    trash_kind = c(colnames(trash_wheel))[7:14],
    mean = c(as.integer(colMeans(trash_wheel[,7:14]))),
    sum = c(as.integer(colSums(trash_wheel[,7:14])))
  ) 

analysis_trash_wheel %>% 
  knitr::kable()
```
The dataset of Mr. Trash Wheel contains **`r nrow(trash_wheel)`** observations. The variables included in the dataset are: `r colnames(trash_wheel)`. During `r min(trash_wheel$date)` to `r max(trash_wheel$date)`, Mr. Trash Wheel collected `r sum(trash_wheel["weight_tons"])` tons of trash. The mean and sum of each kind of trash are displayed in the table above.\

The precipitation dataset after combination covers data in 2018 and 2019 with **`r nrow(precipitation)`** observations. There are three variables: `r colnames(precipitation)`. The sum of all precipitation is **`r sum(precipitation["total"])`**, and the mean of the data is **`r mean(precipitation$total)`**.

Total precipitation in 2018 is **`r sum(precipitation_2018["total"])`**.\
The median number of sports balls in a dumpster in 2019 is **`r median(filter(trash_wheel, year == 2019)$sports_balls)`**.


## Problem 2

Clean the data in pols-month.csv:
```{r}
pols_month = read_csv("fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = month.name[as.numeric(month)],
         president = recode(prez_dem, `0` = "gop", `1` = "dem"),
         year = as.numeric(year)) %>% 
  select(-prez_dem, -prez_gop) %>% 
  select(-day)

pols_month
```
Clean the data in snp.csv:
```{r}
snp = read_csv("fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year>22, year+1900, year + 2000)) %>% 
  select(-day) %>% 
  select(year, month, everything())

snp
```

Clean the data in unemployment.csv:
```{r}
unemployment = read_csv("fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "percentage") %>% 
  mutate(month = as.factor(month)) %>% 
  mutate(month = month.name[month])

unemployment
```
Join the three datasets above:
```{r}
five_thirty_eight = 
  left_join(pols_month, snp, unemployment, by = c("year", "month"))

five_thirty_eight
```

**Conclusion**\
The pols_month dataset has **`r nrow(pols_month)`** rows and **`r ncol(pols_month)`** columns, the variables are `r colnames(pols_month)`. The set covers data from **`r min(pols_month["year"])` to `r max(pols_month["year"])`**. \
The snp dataset contains **`r nrow(snp)`** rows and **`r ncol(snp)`** columns, the variables are `r colnames(snp)`. The set covers data from **`r min(snp["year"])` to `r max(snp["year"])`**. \
The unemployment datset contains **`r nrow(unemployment)`** rows and **`r ncol(unemployment)`** columns, the variables are `r colnames(unemployment)`. The set covers data from **`r min(unemployment["year"])` to `r max(unemployment["year"])`**. \
Joining the three sets by year and month, we get a dataset with **`r nrow(five_thirty_eight)`** rows and **`r ncol(five_thirty_eight)`** columns. The data cover from **`r min(five_thirty_eight["year"])` to `r max(five_thirty_eight["year"])`**. The variables are `r colnames(five_thirty_eight)`.

## Problem 3

Load and clean the data of popularity of baby names:
```{r}
popular_baby_names = read_csv("Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender),
         ethnicity = str_to_lower(ethnicity),
         ethnicity = recode(ethnicity, "asian and paci" = "asian and pacific islander", "black non hisp" = "black non hispanic", "white non hisp" = "white non hispanic"),
         childs_first_name = str_to_lower(childs_first_name),
         childs_first_name = str_to_title(childs_first_name)) %>% 
  distinct(.keep_all = TRUE)

popular_baby_names
```

Produce a table showing rank in popularity of **Olivia** as a female baby name over time:
```{r}
popular_baby_names %>% 
  filter(childs_first_name == "Olivia", gender == "female") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(names_from = "year_of_birth",
              values_from = "rank")
```
Produce a table showing the most popular name among male children over time:
```{r}
popular_baby_names %>% 
  filter(gender == "male", rank == "1") %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  pivot_wider(names_from = "year_of_birth",
              values_from = "childs_first_name")
```
Produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis) for male, white non-hispanic children born in 2016.
```{r}
scatter_df = popular_baby_names %>% 
  filter(gender == "male", ethnicity == "white non hispanic", year_of_birth == 2016) %>% 
  select(childs_first_name, rank, count)

ggplot(scatter_df, aes(x = rank, y = count)) + geom_point() + labs(title = "rank in popularity VS number of children with a name") + theme(plot.title = element_text(face = "bold", hjust = 0.5))
```




